Cartogram Mapping and its Application to Cancer Data Visualisation

### 2019-08-02 review_article.Rmd
added a reference
index of webpages

@@

<!-- Going from choropleth to cartogram 
the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
@GOINO

A cartogram is an alternative map visual. 
An intention distortion is applied to a geographic map base,

A cartogram is a map purposely distorted so that its spatial properties represent quantities
not directly associated with position on the globe. As thematic maps, cartograms emphasize
the distribution of a variable by changing the area (or lengths) of objects on the map.
Our approach maintains continuity, and preserves
many local features of cartographic lines that provide visual clues to the identity of the distorted
objects.
@ACCAC

@CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed map users as map creators. These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’.

-->
<!--
criticisms:
Certain limitations to the standard graphic techniques have also become apparent, however.
They are slow, and do not permit instant retrieval of information. They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
-->

<!--
Urbanisation has pushed people closer together,
-in workplaces and in homes. 
-this issue is not with the size of the administrative areas,
- the increasing urbanisation is contributing to the changing density of population

A particular spatial structure includes
the individuals affected and how they are connected in
communities, as well as the dynamics of these communities
and their organization into larger units. @SAMGIS
-->

<!-- There are many alternative views of the world-->
<!-- Tobler map projections-->
Choropleths may be considered true topological maps, however, if the land mass presented covers enough of the globe, there must be a transformation or distortion to display in 2D. The amount of distortion is related to the distance displayed @GAMP.
The world projections reflect the frequent distortions seen from altering perspectives. Choropleth maps will always be distorted if they cover enough of the globe, as will photographs of the globe from space. Choropleth creation requires choosing a map projection that shows a favourable distortion of the geography for presenting the set of spatial information.
Diagrams that do not specify a projection can be considered to have some unknown projection. <!-- ?? cite tobler.-->

@@

<!-- Why transform?-->
<!-- Common variables used to create cartograms, e.g. population, mortality -->
Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes. Cartograms provide an alternative visualisation method for statistical and geographical information.
The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
Cartograms may be seen as an extension of map transformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].

<!--
@SE Researchers mapping spatial patterns of disease
for small areas are confronted with a number of
challenges, including cartographic design, the
modifiable area unit problem, and also the “small
number” problem, where some places have no,
or few, disease events and/or a low population at
risk of contracting the disease.

>>>>

<!--
A particular spatial structure includes
the individuals affected and how they are connected in
communities, as well as the dynamics of these communities
and their organization into larger units. @SAMGIS
-->

@@

<!-- What is a cartogram? -->
A cartogram alters the map base with the intention of improving the presentation of the statistic of interest. For a single variable of interest, each map area is changed to emphasise the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].

<!-- There are many alternative views of the world-->
<!-- Tobler map projections-->
Choropleths may be considered true topological maps, however, if the land mass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D. The amount of distortion is related to the distance covered by the landmass displayed @GAMP.
World map projections reflect the frequent perspectives used to view the earth. Choropleth maps will always be distorted if they cover enough of the globe, just as photographs of the globe from space. Choropleth creation requires choosing a map projection that shows a favourable distortion of the geography for presenting the set of spatial information.
Diagrams that do not specify a projection can be considered to have some unknown projection. <!-- ??? cite tobler.-->

<!-- Denominators -->
If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed beyond choosing a projection. 
The purposeful distortion of the map space is beneficial when a uniform density of the map base is desired.
When a map base is transformed according to population density, population becomes a uniformly distributed background for the statistic presented [@GOINO].

@@

<!-- Why transform?-->
<!-- Common variables used to create cartograms, e.g. population, mortality -->
Cartograms provide an alternative visualisation method for statistical and geographical information.
The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
Cartograms may be seen as an extension of map transformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].

The transformations required to create cartograms was historically the role of professional cartographers.  
Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’[@CD].

@SE Researchers mapping spatial patterns of disease
for small areas are confronted with a number of
challenges, including cartographic design, the
modifiable area unit problem, and also the “small
number” problem, where some places have no,
or few, disease events and/or a low population at
risk of contracting the disease.
